---
title: 'Lab 11: Introduction to Genome Informatics Lab'
author: "Jordan Laxa"
date: "2023-02-20"
output:
  pdf_document: default
  html_document: default
---
Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes.

```{r}
gene <- read.csv("gene.txt")
library(stringr)
gene[c('number', 'sample', 'genotype', 'expression')] <- str_split_fixed(gene$sample.geno.exp, ' ', 4)
gene <- gene[c('number', 'sample', 'genotype', 'expression')]
gene = subset(gene, select = -c(number))
gene$expression <- as.numeric(gene$expression)
summary(gene)
geneAG <- subset(gene, genotype=="A/G")
length(which(gene$genotype=="A/G"))
summary(geneAG$expression)
geneGG <- subset(gene, genotype=="G/G")
length(which(gene$genotype=="G/G"))
summary(geneGG$expression)
geneAA <- subset(gene, genotype=="A/A")
length(which(gene$genotype=="A/A"))
summary(geneAA$expression)

```

Q13A. The sample size for genotype A/G is 233 and it's median expression level is 25.065. The sample size for genotype G/G is 121 and it's median expression level is 20.074. The sample size for genotype A/A is 108 and it's median expression level is 31.25.

Q14: Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

```{r}
boxplot(geneAA$expression, geneAG$expression, geneGG$expression, xlab = "genotype", ylab = "expression", names = c("AA", "AG", "GG"))
```

Q14A. The median expression value of G/G is 20.074, which is relatively lower than the median expression value of A/A, 31.25. The SNP decreases the expression of ORMDL3.