---
title: "Lab17"
format: pdf
author: Jordan Laxa
---

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

Q1. What column details the total number of people fully vaccinated? 

Q1A. persons_fully_vaccinated

Q2. What column details the Zip code tabulation area? 

Q2A. zip_code_tabulation_area

Q3. What is the earliest date in this dataset?

Q3A. 2021-01-05
```{r}
min(vax[,"as_of_date"])
```

Q4. What is the latest date in this dataset? 

Q4A. 2023-03-07

```{r}
max(vax[,"as_of_date"])
```

```{r}
skimr::skim(vax)
```

Q5. How many numeric columns are in this dataset? 

Q5A. 13

Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

```{r}
nrow(vax[vax$"persons_fully_vaccinated" == "NA",])
```

Q6A. 16621

Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
nrow(vax[vax$"persons_fully_vaccinated" == "NA",]) / nrow(vax[vax$"persons_fully_vaccinated",]) * 100
```

Q7A. 8.27%

Q8. [Optional]: Why might this data be missing?

Q8A. It might be missing because some clinics might have no reported their vaccination results yet.

```{r}
library(lubridate)
today()
```

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

Q9. How many days have passed since the last update of the dataset? 

```{r}
(today() - vax$as_of_date[1]) - (vax$as_of_date[nrow(vax)] - vax$as_of_date[1])
```

Q9A. 6 days

Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(unique(vax[vax$as_of_date, ]))
```

Q10A. There are 18 unique dates.

```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```

```{r}
sd <- vax[vax$county == "San Diego" , ]
nrow(sd)
```

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego" &
                as_of_date > 2022-11-15)

nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

Q11. How many distinct zip codes are listed for San Diego County?

```{r}
length(unique(filter(vax, county == "San Diego")))
```

Q11A. 18

Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
sd[which.max(sd$county == "San Diego"),]
```
Q12A. 91911 has the largest 12+ population.

Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-11-15”?

```{r}
fullsd <- sum(sd$"persons_fully_vaccinated", na.rm = TRUE)
totsd <- sum(sd$"tot_population", na.rm = TRUE)
fullsd / totsd * 100
```

Q13A. 60.46%

Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-11-15”?

```{r}
vacsd <- (sd$"persons_fully_vaccinated")
totpop <- (sd$"tot_population")
hist(vacsd/totpop)
```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
library(ggplot2)
plot <- ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
plot
```

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-11-15")
```

Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```

Q16A. 

```{r}
plot + geom_hline(yintercept=0.7190967, linetype = "dashed")
```

Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”?

Q17A. 
```{r}
ucsdas <- filter(ucsd, as_of_date < "2022-11-15") 
summary(ucsdas$percent_of_population_fully_vaccinated)
```
```{r}
ggplot(ucsdas) + aes(percent_of_population_fully_vaccinated) + geom_histogram(aes(y=..count..),bins=30)
```

Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```

```{r}
vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
Q19A. They are below the average value of 75%.

Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

Q20A.
```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="royalblue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with a population above 36k are shown") +
  geom_hline(yintercept = 0.7190967, linetype="dashed")
```
Q21. How do you feel about traveling for Thanksgiving Break and meeting for in-person class afterwards?

Q21A. I would love to come back for meeting in class, however the quarter would be over by the time break ends :D.
