---
title: "Lab Class 08"
format: pdf
author: Jordan Laxa
---

```{r}
fna.data <- read.csv("C:\\Users\\jorda\\Downloads\\WisconsinCancer.csv")
wisc.df <- data.frame(fna.data, row.names=1)
head(wisc.df)
wisc.data <- wisc.df[,-1]
```

**Q1. How many observations are in this dataset?**

```{r}
nrow(wisc.data)
```

Q1A. There are 569 observations in this data set.

**Q2. How many of the observations have a malignant diagnosis?**

```{r}
diagnosis <- c(wisc.df[,1])
sum(diagnosis == "M")
```

Q2A. There are 212 observations that have a malignant diagnosis.

**Q3. How many variables/features in the data are suffixed with \_mean?**

```{r}
library(stringr)
cn<-colnames(wisc.data, do.NULL = TRUE)
sum(str_count(cn, "_mean"))
```

Q3A. There are 10 variables in the data that are suffixed with "\_mean".

**Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?**

```{r}
colMeans(wisc.data)

apply(wisc.data,2,sd)

wisc.pr <- prcomp(wisc.data, scale=TRUE)

summary(wisc.pr)
```

Q4A. PC1 captures a proportion of 0.4427 of the original variance.

**Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?**

Q5A. 3 principal components are required to describe at least 70% of the original variance in the data.

**Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?**

Q6A. 7 of the principal components are required to describe at least 90% of the original variance in the data.

**Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?**

```{r}
biplot(wisc.pr)
```

Q7A. There are too many variables in this plot, making it very difficult to understand. We will need to shorten the amount of variables to make it easier to digest.

```{r}
diagnosis = as.factor(diagnosis)
plot(wisc.pr$x[, c(1, 2)], col = diagnosis, 
     xlab = "PC1", ylab = "PC2")
```

**Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?**

```{r}
diagnosis = as.factor(diagnosis)
plot(wisc.pr$x[, c(1, 3)], col = diagnosis,
      xlab = "PC1", ylab = "PC3")
```

Q8A. There is a cleaner cut in the first plot because principal component 1 explains more variance in the original data than principal component 3.

```{r}
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis
library(ggplot2)
ggplot(df) + 
    aes(PC1, PC2, col=diagnosis) + geom_point()
```

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

```{r}
pve <- pr.var / sum(pr.var)
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

```{r}
barplot(pve, ylab = "Percent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

**Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr\$rotation\[,1\]) for the feature concave.points_mean?**

```{r}
wisc.pr$rotation[,1]
```

Q9A. -0.26085376

**Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?**

```{r}
pve
4.427203e-01 + 1.897118e-01 + 9.393163e-02 + 9.393163e-02
```

Q10A. 4 principal components is the minimum number required to explain 80% of the variance in the data.

```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method = "complete")
```

**Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?**

```{r}
plot(wisc.hclust)
abline(h=20, col="red", lty=2)
```

Q11A. The clustering model has 4 clusters at a height of 20.

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
table(wisc.hclust.clusters, diagnosis)
```

**Q12. Can you find a better cluster vs diagnoses match by cutting into a different number of clusters between 2 and 10?**

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k = 8)
table(wisc.hclust.clusters, diagnosis)
```

Q12A. With 8 clusters, we can see that there is cluster 2, which contains only malignant diagnoses. This is a better separation than cluster 1 in the 4 cluster example, which had benign and malignant diagnoses.

**Q13. Which method gives your favorite results for the same data.dist dataset? Explain your reasoning.**

```{r}
wisc.hclustS <- hclust(data.dist, method = "single")

plot(wisc.hclustS)

wisc.hclustC <- hclust(data.dist, method = "complete")

plot(wisc.hclustC)

wisc.hclustA <- hclust(data.dist, method = "average")  

plot(wisc.hclustA)

wisc.hclustD <- hclust(data.dist, method = "ward.D2")  

plot(wisc.hclustD)
```

Q13A. I like the complete method of clustering the dataset because it looks cleaner and more neat than the other methods. It also uses more space on the plot, so more connections can be understood.

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:7]), method = "ward.D2")  
plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
table(grps, diagnosis)
plot(wisc.pr$x[,1:2], col=grps)
plot(wisc.pr$x[,1:2], col=diagnosis)
```

```{r}
g <- as.factor(grps)
levels(g)
g <- relevel(g,2)
levels(g)
plot(wisc.pr$x[,1:2], col=g)
```

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:7]), method="ward.D2")
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)
table(wisc.pr.hclust.clusters, diagnosis)
```

**Q15. How well does the newly created model with four clusters separate out the two diagnoses?**

Q15A. The newly created model does a good job of separating out the two diagnoses.

**Q16. How well do the k-means and hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? Again, use the table() function to compare the output of each model (wisc.km\$cluster and wisc.hclust.clusters) with the vector containing the actual diagnoses.**

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
table(wisc.hclust.clusters, diagnosis)
```

Q16A. The hierarchical clustering model I created in the previous section also does a good job in separating the diagnoses.

**Q17. Which of your analysis procedures resulted in a clustering model with the best specificity? How about sensitivity?**

```{r}
## PCA Clustering
## Sensitivity
188 / (188 + 24)
## Specificity
329 / (329 + 28)

## Before PCA Clustering
## Sensitivity
165 / (165 + 5 + 40 + 2)
## Specificity
343 / (343 + 2 + 12)

```

Q17A. The PCA model has a higher sensitivity while the hierarchical clustering model before PCA has a higher specificity.

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

**Q18. Which of these new patients should we prioritize for follow up based on your results?**

Q18A. We should prioritize patient 2 for follow up because it is likely that they have a malignant condition. Furthermore, it is likely that patient 1 has a benign condition.
