---
title: "Class 10: Halloween Mini-Project"
format: html
author: Jordan Laxa
---

```{r}
candy_file <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv")
candy = data.frame(candy_file, row.names=1)
head(candy)

```

Q1. How many different candy types are in this dataset?

```{r}
 nrow(candy)
```
Q1A. There are 85 different candy types in this dataset.

Q2. How many fruity candy types are in the dataset?

```{r}
fruity_candy <- c(candy[,"fruity"])
sum(fruity_candy == "1")
```
Q2A. There are 38 fruit candy types in this dataset.

Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
 candy["Nestle Butterfinger", ]$winpercent
```
Q3A. Neste Butterfinger and 70.74%. 

Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```
Q4A. 76.77%

Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
Q5A. 49.65%

```{r}
library("skimr")
skim(candy)
```

Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Q6A. The sugarpercent, pricepercent, and winpercent columns look to be on a different scale to the majority of the other columns in the data set. The other columns contain only ones and zeroes while these columns contain many decimals.

Q7. What do you think a zero and one represent for the candy$chocolate column?

Q7A. A zero for the chocolate column indicates that the candy does not contain chocolate. A one represents that a candy does contain chocolate.

Q8. Plot a histogram of winpercent values
```{r}
library(ggplot2)
ggplot(candy, aes(x=winpercent)) + geom_histogram()
```

Q9. Is the distribution of winpercent values symmetrical?

Q9A. The distribution of winpercent values is not symmetrical, it's skewed right. 

Q10. Is the center of the distribution above or below 50%?

Q10A. The center of the distribution is below 50%. It is 47.83%.

Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate_mean <- mean(candy$winpercent[as.logical(candy$chocolate)])
fruity_mean <- mean(candy$winpercent[as.logical(candy$fruity)])
```

Q11A. On average, chocolate candy is higher ranked than fruit candy. 

Q12. Is this difference statistically significant?

```{r}
t.test(candy$winpercent[as.logical(candy$chocolate)], candy$winpercent[as.logical(candy$fruity)])
```
Q12A. Yes, this difference is statistically significant because the probability of the t-value being 6.2582 is 2.871e-08, which is extremely small.

Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```
Q13A. The five least liked candy types in this set are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters.

Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),], n=5)
```
Q14A. The top 5 all time favorite candy types are Snickers, Kit Kat, Twix, Reese's Miniatures, and Reese's Peanut Butter cup.

Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)
ggplot(candy) +
  aes(winpercent, rownames(candy)) + geom_col()
```

Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy),winpercent)) + geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

Q17. What is the worst ranked chocolate candy?

Q17A. The worst ranked chocolate candy is Sixlets.

Q18. What is the best ranked fruity candy?
 
Q18A. The best ranked fruity candy is Starburst.

```{r}
library(ggrepel)
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?


```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
tail( candy[ord,c(11,12)], n=5 )
```
Q19A. Tootie Roll Midgies

Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

Q20A. The top 5 most expensive candy types in the dataset are Nik L Nip	
Nestle Smarties	
Ring pop	
Hershey's Krackel	
Hershey's Milk Chocolate. Nik L Nip is the least popular. 

```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Q22A. Chocolate and fruity

Q23. Similarly, what two variables are most positively correlated?

Q23A. Winpercent and chocolate

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[, c(1, 2)], xlab= "PC1", ylab="PC2")
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
library(plotly)
```
```{r}
ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Q24A. Fruity, hard, and pluribus are picked up strongly by PC1 in the positive direction. This makes sense to me because the are characteristics that are usually not shared with chocolate candies. 